---
title: "Introduction to CPLASS"
author: "Linh Do and Scott A. McKinley"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this rmarkdown file, we introduce CPLASS algorithm for detecting changes in velocity.

# Call function and load package
```{r call function, warning=FALSE, message=FALSE}
source("CPLASS support files/CPLASS.R")
source("CPLASS support files/simulation_functions.R")
source("CPLASS support files/plot_functions.R")
source("CPLASS support files/csa_functions.R")
source("CPLASS support files/msd_functions.R")
```

- "CPLASS.R" contains all functions: stochastic search, piecewise linear continuous approximation, and the main CPLASS algorithm.

- "simulation_functions.R" contains simulation functions associated with Cook and et.al. paper on optimal framerate.

- "plot_functions.R" contains all of plot functions for visualization purposes.

- "csa_functions.R" contains functions for calculating cummulative speed allocation

- "msd_function.R" contains functions for calculating mean square displacement

# Run CPLASS on your simulated or real dataset

In this section, we will give an example of how to run CPLASS on your data.

## CPLASS function

### For a single path

**Description**

This function runs the Continuous Piecewise Linear Approximation with Stochastic Search (CPLASS) algorithm on a 2D data in the form of $(x_i,y_i)_{i=1}^n$ observed at time $(t_i)_{i=1}^n$ believed to follow a continuous piecewise linear regression model (Gaussian noise). It is used for detecting changes in velocity problem. CPLASS returns the time changes, the estimated parameters.

**Usage**

<pre> CPLASS(t,x, y,time_rate, lambda_r=1/30, iter_max = 5000, burn_in=500, s_cap=1, gamma=1.2, speed_pen=TRUE) </pre>

**Arguments**

| Argument     | Description |
|--------------|-------------|
| **x**        | A vector containing the data sequence (Cargo locations in x-axis) |
| **y**        | A vector containing the data sequence (Cargo locations in y-axis) |
| **t**        | A vector containing time |
| **time_rate**        | Time step, e.g., 0.001, 0.01, 0.04, 0.05, 0.1, 1 |
| **lambda_r** | the rate used in the proposal of a new vector of changepoints |
| **iter_max** | The maximum number of iterations for running Metropolis-Hastings searching algorithm |
| **burn_in**    | The number of burn-in steps in MH search algorithm|
| **s_cap**    | The threshold for the output speed. If the inferred speed exceed `s_cap` and the `speed_pen` is activated, then the extra speed penalty will be introduced|
| **gamma**      | The power in the strengthened Schwarz Information Criterion (sSIC) |
| **speed_pen**  | If `TRUE`, adding the speed penalty to the penalty function; if `FALSE`, we only use the linear penalty term sSIC |

**Output**

A list of `segment_inferred` and `path_inferred` will be returned after running the algorithm. 

- The `segment_inferred` is a tibble containing 8 columns 

| Columns     | Description |
|--------------|-------------|
| **cp_times** | The inferred change times |
| **durations**| The inferred segment durations |
| **states**   | A binary vector labeling the state of the associated segments, `0` for stationary, `1` for motile. The labels are created using the cut-off method with a threshold of 100nm/s |
| **speeds**   | The inferred segment speeeds |
| **vx** | The inferred velocity with respect to x-axis |
| **vy** | The inferred velocity with respect to y-axis|

- The `path_inferred` is a tibble containing 6 columns

| Columns     | Description |
|--------------|-------------|
| **t** | The input time |
| **j**| The label of time points corresponding to the labels of segments after using the cut-off method |
| **x**   |The observed data with respect to x-axis  |
| **y**   | The observed data with respect to y-axis |
| **a** | The inferred piecewise linear lines (anchor locations w.r.t x-axis) |
| **b** | The inferred piecewise linear lines (anchor locations w.r.t y-axis)|

### For a collection of paths

For running CPLASS on a collection of paths, we introduced the function `CPLASS_paths`.

**Usage**

<pre> CPLASS_paths(data, time_rate, PARALLEL = FALSE, lambda_r=1/30, iter_max = 5000, burn_in=500, s_cap=1, gamma=1.2, speed_pen=TRUE) </pre>

**Arguments**

| Argument     | Description |
|--------------|-------------|
| **data**     | A list of paths where for each path we can specify $t,x,y$ |
| **time_rate**| Time step, e.g., 0.001, 0.01, 0.04, 0.05, 0.1, 1 |
| **PARALLEL** | If `TRUE`, running parallel computing; if `FALSE`, running sequentially
| **lambda_r** | the rate used in the proposal of a new vector of changepoints |
| **iter_max** | The maximum number of iterations for running Metropolis-Hastings searching algorithm |
| **burn_in**    | The number of burn-in steps in MH search algorithm|
| **s_cap**    | The threshold for the output speed. If the inferred speed exceed `s_cap` and the `speed_pen` is activated, then the extra speed penalty will be introduced|
| **gamma**      | The power in the strengthened Schwarz Information Criterion (sSIC) |
| **speed_pen**  | If `TRUE`, adding the speed penalty to the penalty function; if `FALSE`, we only use the linear penalty term sSIC |

**Output**

A list of paths, in each path, there is a similar output with two sublists `segments_inferred` and `path_inferred` as described in the `CPLASS` function.

## Running CPLASS 
```{r load data}
data = readRDS("data_set/CPLASS_21PN.rds") #load your data here
```
### For a single path
```{r running CPLASS}
# Step 1: what does your path look like?
t = data[[1]]$this_t
x = data[[1]]$this_x
y = data[[1]]$this_y
time_rate = 0.05

# Step 2: running CPLASS
cplass = CPLASS(t,x,y,time_rate,iter_max = 2000) #notice that I am using the default but you can change any argument to adjust the MCMC steps, burn-in steps, speed limit threshold, turn-on or turn off the speed penalty, and so on

# Step 3: save the output as .rds file
saveRDS(cplass,file="Example_path_1_Perinuclear.rds")

```
**Visualize the results**

```{r visualize the output path from CPLASS, warning =FALSE}
i=1
motor = "Perinuclear Lysosome" #enter the name you want to display in the figure
speed_max = c()
speed_max = c(speed_max,max(cplass$segments_inferred$speeds))
max_speed = ceiling(max(speed_max))
plot_path_inferred(cplass,motor = motor,max_speed = max_speed,t_lim = c(0,30))
```

### For a collection of paths

In this section, we give an example of how to run CPLASS on a collection of paths. 

We provide 2 options for you to use, either running sequentially or parallel running.
If you want to use `PARALLEL` running, then please set it to be `TRUE`.

The following code using parallel computing

```{r,warning=FALSE}
# We will run on the 21PN data set with 3 paths.
subdata = data[1:3] #contains 3 paths
cplass_paths = CPLASS_paths(subdata, time_rate = 0.05,iter_max = 1000, PARALLEL = TRUE)
saveRDS(cplass_paths,file="Example_3_paths_Perinuclear.rds")
```



**Visualization**

Since we have a lot of paths here, it is better to save the plots into a pdf file to see them all together. The follow code will help us to do so.

```{r visualize the output, warning =FALSE}
path_list = cplass_paths

#######################
# for REAL DATA use the lines below.
# choose which data you are reading (index 1 or 2)
file_list = c("21PF","21PN")
filestub = paste0("CPLASS Final/Real",file_list[2])
# infile = paste0(filestub,"_cplass.rds")
outfile = paste0(filestub,"_gallery.pdf")
# path_list = readRDS(infile)
path_list = cplass_paths
# 
motor = "Perinuclear Lysosomes" #change the name to the name of your data

#######################
# #for SIMULATED DATA use the lines below.
# filestub = paste0("OFR Final/Sim",experiment_num,"_",theta$motor,"_",Hz,"Hz_cplass_",sp)
# infile = paste0(filestub,".rds")
# outfile = paste0(filestub,"_gallery_",sp,".pdf")
# path_list = readRDS(infile)
# 
# if(theta$motor == "CKP"){
#     motor = "Base"
#   }else if(theta$motor == "kin1"){
#     motor = "Contrast"
#   }else if(theta$motor == "Mimic"){
#     motor = "Mimic"
# }


pdf_output = TRUE

if (pdf_output == TRUE){
  pdf(outfile,onefile = TRUE,width=7,height=5)
}

dist_max = c()
t_max = c()
speed_max = c()

for (i in 1:length(path_list)){
  path = path_list[[i]]$path_inferred
  segments = path_list[[i]]$segments_inferred
  speed_max = c(speed_max,max(path_list[[i]]$segments_inferred$speeds))
  
  dist_max = c(dist_max,max(c(max(path$x)-min(path$x)),
                            max((path$y)-min(path$y))))
  t_max = c(t_max,max(path$t) - min(path$t))
}
xy_width = 1.2*max(dist_max)
t_lim = c(0,ceiling(max(t_max)))
max_speed = ceiling(max(speed_max))

for (i in 1:length(path_list)){
  if (i %% 20 == 0){
    print(paste("Working on path",i))
  }
  
  # #REAL DATA PLOTS (dashboard)
  # dashboard = plot_path_real(path_list[[i]],xy_width,t_lim)
  # print(dashboard)

  # SIMULATION PLOTS (dashboard)
  dashboard = plot_path_inferred(path_list[[i]],xy_width,t_lim, motor, max_speed)
  print(dashboard)

  # SIMULATION PLOTS (only x-y plot)
  # plot_xy = plot_path_inferred_xy(path_list[[i]],xy_width,t_lim, motor, max_speed)
  # print(plot_xy)
  
}

if (pdf_output == TRUE){
  dev.off()
}
```
